---
title: "fyp_study2"
author: "Jinxiao Zhang"
date: "January 25, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load library}
# library("knitr")
library("kableExtra") # for making nice tables
library("GGally")
library("lme4")
library("janitor")    # for cleaning column names
library("broom")      # for tidying up linear models
library("tidyverse")
library("cowplot")
library("lubridate") # load lubridate for time data
library("brms")

load("data/workspace_fyp_study2.Rdata")
```

```{r set theme}
theme_set(theme_classic() ) #set the theme
```


```{r participant list}
# (server must be mounted with appropriate file path)
bp_path <- "Z:/SBER/SBER1_physio_R56/data/Full_Study/byparticipant_data"
(bp_files <- list.files(bp_path))

# for FYP purpose
# authenticate
gs_auth()
gs_ls()
# import from Google Sheet
ss = gs_title("Study Quality Control: Data Checking")
gs_data_check <- gs_read(ss, ws = 2, range = cell_rows(3:107))
pids <- gs_data_check$`Participant ID`
```

```{r import raw task data}
# create the list of csv files of ER task
ercsvs <- character(0)
list_noS2 <- c() # list of participants who had no S2 .csv files
for(pid in pids){
  # create list of session 2 csvs
  s2path <- sprintf("%s/%s/session2", bp_path, pid)
  s2ls <- list.files(s2path, recursive = T, full.names = T)
  fn <- s2ls[c(grep("InstructedER_[0-9]{4}.*csv$", s2ls), 
               #grep("Instructed_ER_Task.*csv$", s2ls), 
               grep("Instructed_ER.*csv$", s2ls))]
  if(length(fn) > 0){
    #print(c(pid, length(fn)))
    ercsvs <- append(ercsvs, fn)
  }else{
    list_noS2 = append(list_noS2, pid)
  }
}
# sort the No_session2 list
list_noS2 <- sort(list_noS2)
# check the files in the No_session2 list
for(pid in list_noS2){
  # create list of session 2 csvs
  s2path <- sprintf("%s/%s/session2", bp_path, pid)
  s2ls <- list.files(s2path, recursive = T, full.names = T)
  # print the ID and the number of files in the "session2" folder
  print(c(pid,length(s2ls)))
}
# 1275 - copy (2 identical files) --> remove the copy
# 1574 - pt1, pt2 (2 files both only had 38 rows)
# 5767 - 2 files created at different time
# 1754 - 2 files created at different time
# 1016 - V1, V2, V3 (V1 and V2 are empty files though)
# 1023 - files in a subfolder "Random Files and test data" --> remove the files in the subfolder
# 3538 - 2 files created at different time

# remove the wrong files
ercsvs <- ercsvs[c(-grep("copy", ercsvs), # contain "copy"
                   -grep("Random Files", ercsvs))] # in the "Random Files" folder




#import the raw data
agg_df_raw <- mutate_all(read.csv(ercsvs[1]), as.character) #aggregated raw data
for(i in 2:length(ercsvs)){
  # aggregate all csvs into 1 dataframe
  agg_df_raw <- bind_rows(agg_df_raw, mutate_all(read.csv(ercsvs[i]), as.character))
}

```

```{r fatigue and focus in task}

########## check the fatigue/quality rating in task #########

#get the block-wise rating data
fatigue_quality_rating <- agg_df_raw %>% 
  group_by(participant) %>% 
  filter(imageFile == "") %>% 
  select(participant, 
         Fatigue.response, 
         Data_Quality.response, 
         Watch_Quality.response,
         Distract_Quality.response,
         Reappraise_Quality.response) %>% 
  filter(row_number()!= 1) %>% # remove the trial before the first block
  mutate(fatigue = as.numeric(Fatigue.response),
         focus = as.numeric(Data_Quality.response),
         watch_quality = as.numeric(Watch_Quality.response),
         distract_quality = as.numeric(Distract_Quality.response),
         reappraise_quality = as.numeric(Reappraise_Quality.response),
         block = row_number()) %>% 
  filter(block <= 5)

# plot of ratings by block
# fatigue
p_fatigue <- ggplot(fatigue_quality_rating, aes(fatigue))+
  geom_histogram(colour="black", fill="white")+
  #geom_density(alpha=.2, fill="red") +
  facet_wrap(.~block, ncol = 5)+
  labs(title = "fatigue rating by block")
p_focus <- ggplot(fatigue_quality_rating, aes(focus))+
  geom_histogram(colour="black", fill="white")+
  #geom_density(alpha=.2, fill="red") +
  facet_wrap(.~block, ncol = 5)+
  labs(title = "focus rating by block")
p_watch <- ggplot(fatigue_quality_rating, aes(watch_quality))+
  geom_histogram(colour="black", fill="white")+
  #geom_density(alpha=.2, fill="red") +
  facet_wrap(.~block, ncol = 5)+
  labs(title = "watch rating by block")
p_distract <- ggplot(fatigue_quality_rating, aes(distract_quality))+
  geom_histogram(colour="black", fill="white")+
  #geom_density(alpha=.2, fill="red") +
  facet_wrap(.~block, ncol = 5)+
  labs(title = "distract rating by block")
p_reappraise <- ggplot(fatigue_quality_rating, aes(reappraise_quality))+
  geom_histogram(colour="black", fill="white")+
  #geom_density(alpha=.2, fill="red") +
  facet_wrap(.~block, ncol = 5)+
  labs(title = "reappraise rating by block")

plot_grid(p_fatigue, 
             p_focus, 
             p_watch,
             p_distract,
             p_reappraise,
             ncol=1)

# summarize each participant block ratings
fatigue_quality_summary <- fatigue_quality_rating %>% 
  group_by(participant) %>% 
  summarise(fatigue = mean(fatigue, na.rm = T),
            focus = mean(focus, na.rm = T),
            watch_quality = mean(watch_quality, na.rm = T),
            distract_quality = mean(distract_quality, na.rm = T),
            reappraise_quality = mean(reappraise_quality, na.rm = T),
            n = n()) %>% 
  gather(rating_type, score, fatigue:reappraise_quality)

# add levels of rating types
fatigue_quality_summary$rating_type <- factor(fatigue_quality_summary$rating_type,
                                              levels = c("fatigue",
                                                         "focus",
                                                         "watch_quality",
                                                         "distract_quality",
                                                         "reappraise_quality"))

# plot the distribution
ggplot(fatigue_quality_summary, aes(x = score))+
  geom_histogram(aes(y=..density..), colour="black", fill="white")+
  geom_density(alpha=.2, fill="red") +
  facet_wrap(.~rating_type, ncol = 3)

ggplot(fatigue_quality_summary, aes(x = score))+
  geom_histogram(aes(), colour="black", fill="white")+
  facet_wrap(.~rating_type, ncol = 3)


###### distribution of rating data #####
ggplot(df.ER_ind_wide, aes(watch_val_rating))+
  geom_histogram(aes(y = ..density..), fill = "white", color = "black")+
  geom_density(fill = "salmon", alpha = .5)+
  facet_wrap(. ~ intensity)+
  labs(title = "Distribution of valence rating")

ggsave("plots/ER_val_distribution.png", width = 8, height = 6)

ggplot(df.ER_ind_wide, aes(watch_arou_rating))+
  geom_histogram(aes(y = ..density..), fill = "white", color = "black")+
  geom_density(fill = "salmon", alpha = .5)+
  facet_wrap(. ~ intensity)+
  labs(title = "Distribution of arousal rating")

ggsave("plots/ER_arou_distribution.png", width = 8, height = 6)

```
```{r create ER task individual-level data}
# create a clean data from the raw data
agg_df <-  agg_df_raw
agg_df$task <- tolower(agg_df$task) #change it to lower case
agg_df$task <- gsub("rethink", "reappraise", agg_df$task) #replace "rethink" with reappraise
#remove unnecessary rows of "instructions"
agg_df <- agg_df %>% 
  filter(imageFile != "")
#force variable type to be numeric
agg_df$NegRating.response <- as.numeric(agg_df$NegRating.response)
agg_df$ArousalRating.response <- as.numeric(agg_df$ArousalRating.response)

# remove participants who had less than 100 trials
agg_df <- agg_df %>% 
  group_by(participant) %>% 
  mutate(n_trial = n()) %>% 
  filter(n_trial > 100)
# change the variable "assign" to "intensity"
names(agg_df)[names(agg_df) == "assign"] <- "intensity"

# check the number of neutral, low, mid, and high trials (trial number = 15 for each condition)
agg_df %>% 
  group_by(participant, task, intensity) %>% 
  tally()
# write the data to the server
write_csv(agg_df, "Z:/SBER/SBER1_physio_R56/data/Full_Study/session2_data/ier/IER_trial_level.csv")

# individual-wise data
df.ER_ind <-  agg_df %>% 
  group_by(participant, task, intensity) %>% 
  summarise(NegRating.response = mean(NegRating.response, na.rm = TRUE),
            ArousalRating.response = mean(ArousalRating.response, na.rm = TRUE)) %>% 
  filter(!is.nan(NegRating.response), !is.nan(ArousalRating.response)) # some participants have no response and should be removed, e.g. 3032

df.ER_ind$task <- factor(df.ER_ind$task, levels=c('watch', 'reappraise','distract'), ordered=TRUE)
df.ER_ind$intensity <- factor(df.ER_ind$intensity, levels=c('high', 'mid','low', 'ntr'), ordered=TRUE)


################### change scores ####################
#valence ratings changes
temp_val_change <-  df.ER_ind %>% 
  spread(task, NegRating.response) %>% 
  group_by(participant, intensity) %>% 
  summarise(watch_val_rating = sum(watch, na.rm = T), 
            distract_val_rating = sum(distract, na.rm = T), 
            reappraise_val_rating = sum(reappraise, na.rm = T))
#arousal ratings changes
temp_arou_change <-  df.ER_ind %>% 
  spread(task, ArousalRating.response) %>% 
  group_by(participant, intensity) %>% 
  summarise(watch_arou_rating = sum(watch, na.rm = T), 
            distract_arou_rating = sum(distract, na.rm = T), 
            reappraise_arou_rating = sum(reappraise, na.rm = T))
# merge the 2 wide dataframes
df.ER_ind_wide <- merge(temp_val_change, temp_arou_change, 
                       by = c("participant","intensity"))
df.ER_ind_wide <- df.ER_ind_wide %>% 
  mutate(distract_val_change = watch_val_rating - distract_val_rating, 
         reappraise_val_change = watch_val_rating - reappraise_val_rating,
         distract_arou_change = watch_arou_rating - distract_arou_rating, 
         reappraise_arou_change = watch_arou_rating - reappraise_arou_rating,
         picture = ifelse(intensity == "ntr", "ntr", "neg")) #label the trials as eigher "neutral" or "negative

# write the data to the server
write_csv(df.ER_ind_wide,
          "Z:/SBER/SBER1_physio_R56/data/Full_Study/session2_data/ier/IER_participant_level.csv")
```


```{r check psg files}
linda_folder <- "Z:/SBER/SBER1_physio_R56/SBER_sleepreports_LindaClete/Scored/11.13.18"
destination_folder <- "Z:/SBER/SBER1_physio_R56/data/brux_scored"

dir_list_scored <- list.dirs(destination_folder)
sleep_dirs <- grep("[0-9]{4}_BP\\.SLP$", dir_list_scored, value = T)

```


```{r PSG data processing}
#report path: Z:\SBER\SBER1_physio_R56\SBER_sleepreports_LindaClete\Scored\11.13.18\ScoredPsgs_Reports\09.27.18
#hynogram export path: Z:\SBER\SBER1_physio_R56\data\brux_scored\raw_exports

# the path of raw PSG data
psg_path = "Z:/SBER/SBER1_physio_R56/data/brux_scored/raw_exports"
# list and select the hypnogram data files
psg_files <- list.files(psg_path, full.names = T)
psg_txt_files <- psg_files[grep("hypno.TXT$", psg_files)]
# get the participant numbers
psg_pids <-  str_sub(psg_txt_files, 
                     str_length(psg_txt_files)-13, 
                     str_length(psg_txt_files)-10)

#### read files and calculate the sleep staging information #####
# the raw psg data fram
df.psg <- data.frame()
# the data frame of psg summary report
df.psg_summary <-  data.frame()

for (j in 1:length(psg_pids)){
  # read a single psg file
  psg <- read.csv(psg_txt_files[j]) %>% 
    rename("stage" = "X.") %>% 
    mutate(pid = psg_pids[j],
           time = row_number()/2) %>% # t - time in minute
    select(pid, stage, time)

  
  # form some cases: replace 1, 2, 3 with N1, N2, N3
  psg$stage = as.character(psg$stage)
  psg$stage[psg$stage == "1"] <-  "N1"
  psg$stage[psg$stage == "2"] <-  "N2"
  psg$stage[psg$stage == "3"] <-  "N3"
  
  # combine the single psg file to the whole data frame
  df.psg = rbind(df.psg, psg)
  
  
  
  
  #### calculate the PSG report for each participant #######
  # time point of lights out
  time_light_out = 1
  for (i in 1:nrow(psg)){
    if (psg$stage[i] != "?"){
      time_light_out = i/2 
      break
    } 
  }
  # time point of sleep onset
  time_slp_onset = 1
  for (i in (time_light_out*2):nrow(psg)){
    if (psg$stage[i] %in% c("N1", "N2", "N3")){
      time_slp_onset = i/2 
      break
    } 
  }
  # time point of REM onset
  time_rem_onset = 1
  for (i in (time_slp_onset*2):nrow(psg)){
    if(psg$stage[i] == "R"){
      time_rem_onset = i/2 
      break
    }
  }
  # time point of lights on
  time_light_on = 1
  for (i in 1:nrow(psg)){
    if (psg$stage[nrow(psg) +1 - i] != "?"){
      time_light_on = (nrow(psg) +2 - i)/2 
      break
    } 
  }
  # time point of sleep offset
  time_slp_offset = 1
  for (i in (nrow(psg) - time_light_on*2):nrow(psg)){
    if (psg$stage[nrow(psg) +1 - i] %in% c("N1", "N2", "N3", "R")){
      time_slp_offset = (nrow(psg) +2 - i)/2 
      break
    } 
  }
    # time point of lights out
  time_light_out = 1
  for (i in 1:nrow(psg)){
    if (psg$stage[i] != "?"){
      time_light_out = i/2 
      break
    } 
  }
  # total recording time
  psg_trt = nrow(psg)/2
  # end recording time (recording after lights on)
  psg_ert = psg_trt - time_slp_offset
  # time in bed (time available for sleep)
  psg_tib = time_light_on - time_light_out
  # total sleep time
  psg_tst = psg %>% filter(stage %in% c("N1", "N2", "N3", "R")) %>% nrow()/2
  # WASO - wake after sleep onset
  psg_waso = psg %>% 
    filter(time >= time_slp_onset, time < time_light_on) %>% 
    filter(stage == "W") %>% 
    nrow()/2
  # sleep efficiency
  psg_slp_eff = psg_tst/psg_tib
  # sleep onset latency
  psg_slp_latency = time_slp_onset - time_light_out
  # REM latency
  psg_rem_latency = time_rem_onset - time_slp_onset
  # during the sleep period
  psg_sleep <- psg %>% 
    filter(time >= time_slp_onset, time < time_slp_offset) %>% 
    group_by(stage) %>% 
    summarise(duration = n()/2) %>% 
    spread(stage, duration)
  
  # wake during sleep period
  psg_wake = ifelse(is.null(psg_sleep$W[1]), 0, psg_sleep$W[1])
  # N1 sleep
  psg_n1 = ifelse(is.null(psg_sleep$N1[1]), 0, psg_sleep$N1[1])
  psg_n1_perc = psg_n1/psg_tst
  # N2 sleep
  psg_n2 = ifelse(is.null(psg_sleep$N2[1]), 0, psg_sleep$N2[1])
  psg_n2_perc = psg_n2/psg_tst
  # N3 sleep
  psg_n3 = ifelse(is.null(psg_sleep$N3[1]), 0, psg_sleep$N3[1])
  psg_n3_perc = psg_n3/psg_tst
  # REM sleep
  psg_rem = ifelse(is.null(psg_sleep$R[1]), 0, psg_sleep$R[1])
  psg_rem_perc = psg_rem/psg_tst
  
  # contrust the PSG summary for one participant
  psg_summary <- c(psg_pids[j], psg_trt, psg_ert, psg_tib, psg_tst, psg_waso, psg_slp_eff, 
                   psg_slp_latency, psg_rem_latency, psg_wake, psg_n1, psg_n1_perc,
                   psg_n2, psg_n2_perc, psg_n3, psg_n3_perc, psg_rem, psg_rem_perc) %>% 
    as.numeric()
  # combine the single summary to the whoe summary data frame
  df.psg_summary <- rbind(df.psg_summary, psg_summary)
  
  # set variable names of the summary report
  if(j == 1){
    names(df.psg_summary) <- c("pid", "psg_trt", "psg_ert", "psg_tib", "psg_tst", "psg_waso",
                               "psg_slp_eff", "psg_slp_latency", "psg_rem_latency",
                               "psg_wake", "psg_n1", "psg_n1_perc", "psg_n2",
                               "psg_n2_perc", "psg_sws", "psg_sws_perc", 
                               "psg_rem", "psg_rem_perc")
  }
  
}

# remove psg summary of participants who only have "?" data points
for (i in 1:nrow(df.psg_summary)){
  if(df.psg_summary$psg_tst[i] == 0) {
    # if their total sleep time = 0 (all data points are ?)
    df.psg_summary[i, 3:17] <- NA # assign NA to all summary statistics except the pid and trt
  }
}
df.psg_summary$pid <- as.character(df.psg_summary$pid)

# check participants with NA data
# df.psg_summary %>% filter(is.na(psg_tst)) %>% select(pid)
 


# import the redcap biocalibration log
df.biocali_redcap <- read_csv("Z:/SBER/SBER1_physio_R56/data/Full_Study/redcap/Biocalibration_DATA_2019-03-08_0956.csv") %>% 
  filter(!participant_id %in% c("3694", "5811"))  # remove the first log of 3694 and 5811  (they have a 2nd psg)

df.biocali_redcap <- df.biocali_redcap %>% 
  mutate(pid = str_sub(participant_id, 1, 4),
         pid = ifelse(is.na(as.numeric(subject_id)), pid,
           ifelse(as.numeric(subject_id) > as.numeric(pid), subject_id, pid)),# e.g. 1070
         lack_colon = (!is.na(recording_start_time_pst) & 
                                !grepl(":", recording_start_time_pst)),# decide which cell lacks ":" in the time
        # insert a ":"
         recording_start_time_pst_corrected = 
          ifelse(lack_colon == TRUE,
                 str_c(str_sub(recording_start_time_pst, 1, 2), ":",
                       str_sub(recording_start_time_pst, 3, 4), sep = ""),
                                                     recording_start_time_pst)) 


# merge the psg-recording-start-time to psg data
tmp <- df.psg_summary %>% 
  left_join(df.biocali_redcap %>% select(pid, recording_start_time_pst_corrected),
            by = "pid") %>% 
  mutate(
    psg_record_start_hour = hour(strptime(recording_start_time_pst_corrected, "%H:%M")), #get the hour of the recording start time
    psg_record_start = ifelse(
      psg_record_start_hour <= 12, # is the time in the evening?
      (strptime(recording_start_time_pst_corrected, "%H:%M") + hours(12)) %>%
        strftime("%H:%M"),# make corrections if the time is in the morning
      strptime(recording_start_time_pst_corrected, "%H:%M")%>% 
        strftime("%H:%M")),
    psg_record_stop = (strptime(psg_record_start, "%H:%M") + 
                         minutes(as.integer(psg_trt))) # have to be integer for time calculation
    %>% strftime("%H:%M"))

# to check the stop-recording time
tmp %>% arrange(psg_ert) %>% view()


## problematic psg data:
#   5029 - battery died at 3:49
#   5675 - took the psg off at 3 am
## recording time < 8 hours
#   1046 - 275.5; psg stopped recording at 2:32, stopped in REM
#   3694 - 344.5; psg stopped recording at 3:27, stopped in N2
#   2001 - 472.5; reported to go to bed at 10:30 and wake up at 6:30, stop-recording time unknown (around 4:30 am), stopped in REM
#   3741 - 246.5; psg stopped recording at 2:39, stopped in N2
## total sleep time < 200 minutes
#   1070 - 170; took off the psg around 4 am; psg stopped recording at 11:07;
#   1570 - 191.5; reported to go to bed around 12, have 9 hours sleep and waked up at 9:24 on 5/24 - may have taken the psg off - 2 single N2, N3 points way ahead sleep time???


## other no end-recording-time cases (should be fine, keep these)
#   1581 - no end-recording time; go to bed at 11:30, stopped recording around 8:52; reported waked up at 8:50 and got up at 9 - should be fine
#   5461 - no end-recording time; try to sleep at 9, stopped recording around 4:46; reported waked up at 4:55 and got up at 6:15 - should be fine

# assign tmp to df.psg_summary
df.psg_summary <- tmp

# remove problematic psg data
df.psg_summary <- df.psg_summary %>%
  filter(!pid %in% c("5029", "5675", "1046", "3694", "2001", "1070", "1570", "3741"))

# mark the labels for stage
df.psg$stage <- factor(df.psg$stage, 
                         levels = c("?", "N1", "N2", "N3", "R", "W"),
                         labels = c("?", "N1", "N2", "SWS", "R", "W"))

# # plot the staging data to check
# ggplot(df.psg %>% filter(pid == "1570"), aes(time, stage))+
#   geom_point(alpha = .2, size = 1)

# kable(df.psg_summary, digit = 3)

# write the psg dataset
write_csv(df.psg_summary, "Z:/SBER/SBER1_physio_R56/data/Full_Study/psg_data/psg_summary.csv")


```

```{r combine PSG and ER task data}
glimpse(df.psg_summary)
glimpse(df.ER_ind_wide)

# combine ER task data and PSG data
df.er_psg <- left_join(df.ER_ind_wide, df.psg_summary, 
                       by = c("participant" = "pid"))

# read individual difference data
df.ind_diff <- read_csv("Z:/SBER/SBER1_physio_R56/data/Full_Study/inddiff_Qualtrics/raw/SBER1_inddiff_December_3_2018_20.54.csv")

# add gender and age
df.er_psg <- left_join(df.er_psg, df.ind_diff %>% filter(Progress == 100)
                         %>% select(partID, age, gender),
                       by = c("participant" = "partID"))
df.er_psg$age <- as.numeric(df.er_psg$age)

# check how many participants in df.er_psg have complete data
df.er_psg %>% 
  filter(!is.na(psg_tst)) %>%
  select(participant) %>% 
  unique()

# check wparticipants who have psg data but not session 2 ER data
df.psg_summary$pid[!df.psg_summary$pid %in% df.ER_ind_wide$participant]
# 1061 - only 75 trials in ER task
# 1070 - no session 2 data
# 1152 - only 76 trials in ER task
# 1574 - 2 data files, did the first block twice
# 3026 - no IER task data
# 3032 - no response in the task data
# 3036 - no IER task data
```

```{r visualization 1 PSG}
# hist(df.psg_summary$psg_tib, main = "PSG time in bed")
# hist(df.psg_summary$psg_tst, main = "PSG total sleep time")
# hist(df.psg_summary$psg_sws, main = "PSG SWS time")
# hist(df.psg_summary$psg_rem, main = "PSG REM sleep time")
# hist(df.psg_summary$psg_sws_perc, main = "PSG SWS percentage")
# hist(df.psg_summary$psg_rem_perc, main = "PSG REM percentage")
# 
# # REM and SWS
# p_psg1 <- ggplot(df.psg_summary,
#        aes(psg_rem, psg_sws))+
#   geom_point(alpha = .8)+
#   geom_smooth(method = "lm")+
#   labs(title = "REM ~ SWS (min)")
# 
# p_psg2 <- ggplot(df.psg_summary,
#        aes(psg_rem_perc, psg_sws_perc))+
#   geom_point(alpha = .8)+
#   geom_smooth(method = "lm")+
#   labs(title = "REM ~ SWS (%)")
# 
# # TST and REM
# p_psg3 <- ggplot(df.psg_summary,
#        aes(psg_rem, psg_tst))+
#   geom_point(alpha = .8)+
#   geom_smooth(method = "lm")+
#   labs(title = "REM ~ TST (min)")
# 
# # TST and REM
# p_psg4 <- ggplot(df.psg_summary,
#        aes(psg_sws, psg_tst))+
#   geom_point(alpha = .8)+
#   geom_smooth(method = "lm")+
#   labs(title = "SWS ~ TST (min)")


## correlation matrix ####
ggpairs(df.psg_summary %>%
          # filter(intensity == "high") %>% 
          select(psg_trt, psg_tib, psg_tst, 
                 psg_n1, psg_n2, psg_sws, psg_rem
                 # ,watch_val_rating, watch_arou_rating
                 )) +
  theme_grey()


ggsave("plots/psg_correlations.png", width = 10, height = 9)
# ggsave("plots/psg_reactivity_correlations.png", width = 12, height = 9)

# library(corrplot)
# M <- cor(df.er_psg %>% 
#            filter(intensity == "high") %>% 
#            select(psg_tst, psg_n1, psg_n2, psg_n3, psg_rem) %>% 
#            na.omit())
# corrplot.mixed(M, upper = "ellipse")

```
```{r visualization 2 REM and reactivity}
#### REM and valence-reactivity ####
p_rem1 <- ggplot(df.er_psg,
       aes(psg_rem, watch_val_rating, color = intensity))+
  geom_point(alpha = .8)+
  geom_smooth(method = "lm")+
  labs(title = "REM and valence-reactivity (all cases)")

p_rem3 <- ggplot(df.er_psg,
       aes(psg_rem_perc, watch_val_rating, color = intensity))+
  geom_point(alpha = .8)+
  geom_smooth(method = "lm")+
  labs(title = "REM% and valence-reactivity (all cases)")


#### REM and arousal-reactivity ####
p_rem5 <- ggplot(df.er_psg,
       aes(psg_rem, watch_arou_rating, color = intensity))+
  geom_point(alpha = .8)+
  geom_smooth(method = "lm")+
  labs(title = "REM and arousal-reactivity (all cases)")

p_rem7 <- ggplot(df.er_psg,
       aes(psg_rem_perc, watch_arou_rating, color = intensity))+
  geom_point(alpha = .8)+
  geom_smooth(method = "lm")+
  labs(title = "REM% and arousal-reactivity (all cases)")


plot_grid(p_rem1, p_rem5, p_rem3, p_rem7, ncol = 2)
ggsave("plots/rem-reactivity correlation.png", width = 12, height = 9)


```
```{r visualization 3 REM and regulation}
#### REM and regulation of valence ####
p_rem9 <- ggplot(df.er_psg,
       aes(psg_rem, reappraise_val_change, color = intensity))+
  geom_point(alpha = .8)+
  geom_smooth(method = "lm")+
  labs(title = "REM and reappraisal-valence change")

p_rem10 <- ggplot(df.er_psg,
       aes(psg_rem, distract_val_change, color = intensity))+
  geom_point(alpha = .8)+
  geom_smooth(method = "lm")+
  labs(title = "REM and distraction-valence change")

p_rem13 <- ggplot(df.er_psg,
       aes(psg_rem_perc, reappraise_val_change, color = intensity))+
  geom_point(alpha = .8)+
  geom_smooth(method = "lm")+
  labs(title = "REM% and reappraisal-valence change")

p_rem14 <- ggplot(df.er_psg,
       aes(psg_rem_perc, distract_val_change, color = intensity))+
  geom_point(alpha = .8)+
  geom_smooth(method = "lm")+
  labs(title = "REM% and distraction-valence change")


plot_grid(p_rem9, p_rem10, p_rem13, p_rem14)
ggsave("plots/rem-valence-regulation correlation.png", width = 12, height = 9)

#### REM and regulation of arousal ####

p_rem11 <- ggplot(df.er_psg,
       aes(psg_rem, reappraise_arou_change, color = intensity))+
  geom_point(alpha = .8)+
  geom_smooth(method = "lm")+
  labs(title = "REM and reappraisal-arousal change")

p_rem12 <- ggplot(df.er_psg,
       aes(psg_rem, distract_arou_change, color = intensity))+
  geom_point(alpha = .8)+
  geom_smooth(method = "lm")+
  labs(title = "REM and distraction-arousal change (REM > 0)")

p_rem15 <- ggplot(df.er_psg,
       aes(psg_rem_perc, reappraise_arou_change, color = intensity))+
  geom_point(alpha = .8)+
  geom_smooth(method = "lm")+
  labs(title = "REM% and reappraisal-arousal change")

p_rem16 <- ggplot(df.er_psg,
       aes(psg_rem_perc, distract_arou_change, color = intensity))+
  geom_point(alpha = .8)+
  geom_smooth(method = "lm")+
  labs(title = "REM% and distraction-arousal change")

plot_grid(p_rem11, p_rem12, p_rem15, p_rem16)
ggsave("plots/rem-arousal-regulation correlation.png", width = 12, height = 9)
```

```{r visualization 4 SWS and reactivity}
#### SWS and valence-reactivity ####
p_sws1 <- ggplot(df.er_psg,
       aes(psg_sws, watch_val_rating, color = intensity))+
  geom_point(alpha = .8)+
  geom_smooth(method = "lm")+
  labs(title = "SWS and valence-reactivity")

p_sws2 <- ggplot(df.er_psg ,
       aes(psg_sws_perc, watch_val_rating, color = intensity))+
  geom_point(alpha = .8)+
  geom_smooth(method = "lm")+
  labs(title = "SWS% and valence-reactivity")

#### SWS and arousal-reactivity ####
p_sws3 <- ggplot(df.er_psg,
       aes(psg_sws, watch_arou_rating, color = intensity))+
  geom_point(alpha = .8)+
  geom_smooth(method = "lm")+
  labs(title = "SWS and arousal-reactivity")

p_sws4 <- ggplot(df.er_psg,
       aes(psg_sws_perc, watch_arou_rating, color = intensity))+
  geom_point(alpha = .8)+
  geom_smooth(method = "lm")+
  labs(title = "SWS% and arousal-reactivity")

plot_grid(p_sws1, p_sws3, p_sws2, p_sws4)
ggsave("plots/sws-reactivity correlation.png", width = 12, height = 9)
```
```{r visualization 5 SWS and regulation}
#### SWS and valence-regulation ####
p_sws5 <- ggplot(df.er_psg,
       aes(psg_sws, reappraise_val_change, color = intensity))+
  geom_point(alpha = .8)+
  geom_smooth(method = "lm")+
  labs(title = "SWS and reappraisal-valence change")

p_sws6 <- ggplot(df.er_psg,
       aes(psg_sws, distract_val_change, color = intensity))+
  geom_point(alpha = .8)+
  geom_smooth(method = "lm")+
  labs(title = "SWS and distract-valence change")

p_sws7 <- ggplot(df.er_psg,
       aes(psg_sws_perc, reappraise_val_change, color = intensity))+
  geom_point(alpha = .8)+
  geom_smooth(method = "lm")+
  labs(title = "SWS% and reappraisal-valence change")

p_sws8 <- ggplot(df.er_psg ,
       aes(psg_sws_perc, distract_val_change, color = intensity))+
  geom_point(alpha = .8)+
  geom_smooth(method = "lm")+
  labs(title = "SWS% and distract-valence change")

plot_grid(p_sws5, p_sws6, p_sws7, p_sws8)
ggsave("plots/sws-valence-regulation correlation.png", width = 12, height = 9)

#### SWS and arousal-regulation ####
p_sws9 <- ggplot(df.er_psg ,
       aes(psg_sws, reappraise_arou_change, color = intensity))+
  geom_point(alpha = .8)+
  geom_smooth(method = "lm")+
  labs(title = "SWS and reappraisal-arousal change")

p_sws10 <- ggplot(df.er_psg,
       aes(psg_sws, distract_arou_change, color = intensity))+
  geom_point(alpha = .8)+
  geom_smooth(method = "lm")+
  labs(title = "SWS and distract-arousal change")

p_sws11 <- ggplot(df.er_psg ,
       aes(psg_sws_perc, reappraise_arou_change, color = intensity))+
  geom_point(alpha = .8)+
  geom_smooth(method = "lm")+
  labs(title = "SWS% and reappraisal-arousal change")

p_sws12 <- ggplot(df.er_psg,
       aes(psg_sws_perc, distract_arou_change, color = intensity))+
  geom_point(alpha = .8)+
  geom_smooth(method = "lm")+
  labs(title = "SWS% and distract-arousal change")

plot_grid(p_sws9, p_sws10, p_sws11, p_sws12)
ggsave("plots/sws-arousal-regulation correlation.png", width = 12, height = 9)

```

```{r visualization 6 TST}
### TST (total sleep time) and reactivity ####
p_tst1 <- ggplot(df.er_psg,
       aes(psg_tst, watch_val_rating, color = intensity))+
  geom_point(alpha = .8)+
  # geom_text(data = df.er_psg %>% filter(intensity == "high"), 
  #           aes(label = participant), color = "black")+
  geom_smooth(method = "lm", se = F)+
  labs(title = "Total sleep time and valence-reactivity")

# p_tst1b <- ggplot(df.er_psg,
#        aes(psg_tst, watch_val_rating, color = intensity))+
#   geom_point(alpha = .8)+
#   geom_smooth(method = "lm")+
#   facet_wrap(.~intensity, nrow = 1)+
#   labs(title = "Total sleep time and valence-reactivity")

p_tst2 <- ggplot(df.er_psg,
       aes(psg_tst, watch_arou_rating, color = intensity))+
  geom_point(alpha = .8)+
  geom_smooth(method = "lm", se = F)+
  labs(title = "Total sleep time and arousal-reactivity")

# p_tst2b <- ggplot(df.er_psg,
#        aes(psg_tst, watch_arou_rating, color = intensity))+
#   geom_point(alpha = .8)+
#   geom_smooth(method = "lm")+
#   facet_wrap(.~intensity, nrow = 1)+
#   labs(title = "Total sleep time and arousal-reactivity")
# 
# plot_grid(p_tst1b, p_tst2b, ncol = 1)
# ggsave("plots/tst-reactivity_separated.png", width = 12, height = 7)

### TST and regulation ####
p_tst3 <- ggplot(df.er_psg,
       aes(psg_tst, reappraise_val_change, color = intensity))+
  geom_point(alpha = .8)+
  geom_smooth(method = "lm", se = F)+
  labs(title = "Total sleep time and reappraisal-valence change")

p_tst4 <- ggplot(df.er_psg,
       aes(psg_tst, distract_val_change, color = intensity))+
  geom_point(alpha = .8)+
  geom_smooth(method = "lm", se = F)+
  labs(title = "Total sleep time and distraction-valence change")

p_tst5 <- ggplot(df.er_psg,
       aes(psg_tst, reappraise_arou_change, color = intensity))+
  geom_point(alpha = .8)+
  geom_smooth(method = "lm", se = F)+
  labs(title = "Total sleep time and reappraisal-arousal change")

p_tst6 <- ggplot(df.er_psg,
       aes(psg_tst, distract_arou_change, color = intensity))+
  geom_point(alpha = .8)+
  geom_smooth(method = "lm", se = F)+
  labs(title = "Total sleep time and distraction-arousal change")

plot_grid(p_tst1, p_tst2, p_tst3, p_tst4, p_tst5, p_tst6, ncol = 2)
ggsave("plots/tst-reactivity and regulation correlation.png", width = 12, height = 13.5)

```

```{r visualization 7: mean-center the raw rating}

ggplot(df.er_psg,
       aes(psg_tst, watch_arou_rating, color = intensity))+
  geom_point(alpha = .8)+
  geom_smooth(method = "lm", se = F)+
  geom_smooth(data = df.er_psg,
              aes(psg_tst, watch_arou_rating),
              linetype = 2,
              color = "black",
              method = "lm", se = F)+
  labs(title = "Total sleep time and arousal-reactivity (original)")


ggplot(df.er_psg,
       aes(psg_tst, watch_arou_rating))+
  geom_point(aes(color = intensity), alpha = .8)+
  geom_smooth(data = df.er_psg,
              aes(psg_tst, watch_arou_rating),
              linetype = 2,
              color = "black",
              method = "lm", se = F)+
  labs(title = "Total sleep time and arousal-reactivity (average)")


df.er_psg <- df.er_psg %>% 
  group_by(participant) %>% 
  mutate(watch_arou_rating_ctrd = scale(watch_arou_rating, 
                                        center=T, scale = F),
         watch_arou_rating_scld = scale(watch_arou_rating, 
                                        center=T, scale = T))

ggplot(df.er_psg,
       aes(psg_tst, watch_arou_rating_ctrd, color = intensity))+
  geom_point(alpha = .8)+
  geom_smooth(method = "lm", se = F)+
  geom_smooth(aes(psg_tst, watch_arou_rating_ctrd),
              linetype = 2,
              color = "black",
              method = "lm", se = F)+
  labs(title = "Total sleep time and arousal-reactivity (mean-centered)")


```



```{r analysis: PSG and ER-task correlation}
# REM duration
with(df.er_psg %>% filter(intensity == "low"),
     cor.test(psg_rem, watch_val_rating))
with(df.er_psg %>% filter(intensity == "low"),
     cor.test(psg_rem, watch_arou_rating))
# REM percentage %
with(df.er_psg %>% filter(intensity == "high", psg_rem_perc > 0),
     cor.test(psg_rem_perc, watch_val_rating))
with(df.er_psg %>% filter(intensity == "high", psg_rem_perc > 0),
     cor.test(psg_rem_perc, watch_arou_rating))


# SWS duration
with(df.er_psg %>% filter(intensity == "low"),
     cor.test(psg_sws, watch_val_rating))
with(df.er_psg %>% filter(intensity == "low"),
     cor.test(psg_sws, watch_arou_rating))
# SWS percentage %
with(df.er_psg %>% filter(intensity == "high", psg_rem_perc > 0),
     cor.test(psg_sws_perc, watch_val_rating))
with(df.er_psg %>% filter(intensity == "high", psg_rem_perc > 0),
     cor.test(psg_sws_perc, watch_arou_rating))

# TST duration
with(df.er_psg %>% filter(intensity == "high"),
     cor.test(psg_tst, watch_val_rating))
with(df.er_psg %>% filter(intensity == "high"),
     cor.test(psg_tst, watch_arou_rating))

with(df.er_psg %>% filter(intensity == "mid"),
     cor.test(psg_tst, watch_val_rating))
with(df.er_psg %>% filter(intensity == "mid"),
     cor.test(psg_tst, watch_arou_rating))

with(df.er_psg %>% filter(intensity == "low"),
     cor.test(psg_tst, watch_val_rating))
with(df.er_psg %>% filter(intensity == "low"),
     cor.test(psg_tst, watch_arou_rating))

# TST and regulation
with(df.er_psg %>% filter(intensity == "mid", psg_rem > 0),
     cor.test(psg_tst, reappraise_arou_change))
```
```{r analysis: REM and ER-task linear models}
## DV: valence rating ####
mod_tst1 <- lm(watch_val_rating ~ psg_tst, 
               data = df.er_psg %>% filter(intensity == "high" , psg_rem > 0))
summary(mod_tst1)

mod_rem1 <- lm(watch_val_rating ~ psg_rem, 
               data = df.er_psg %>% filter(intensity == "high" , psg_rem > 0))
summary(mod_rem1)

mod_rem2 <- lm(watch_val_rating ~ psg_tst + psg_rem, 
               data = df.er_psg %>% filter(intensity == "high" , psg_rem > 0))
summary(mod_rem2)

# The Bayes Factor way
mod_rem2_BF <- regressionBF(watch_val_rating ~ psg_tst + psg_rem, 
               data = df.er_psg %>% filter(intensity == "high") %>% na.omit())
plot(mod_rem2_BF)

## interaction effect

# recode the intensity into order ranking
df.er_psg <- df.er_psg %>% 
  mutate(intensity_num = ifelse(intensity == "ntr", 0,
                                ifelse(intensity == "low", 1,
                                       ifelse(intensity == "mid", 2,
                                              ifelse(intensity == "high", 3,
                                                     NA)))))

mod_rem3 <- lmer(watch_val_rating ~ psg_rem+intensity_num + (1|participant), 
               data = df.er_psg )
mod_rem4 <- lmer(watch_val_rating ~ psg_rem*intensity_num + (1|participant), 
               data = df.er_psg)
anova(mod_rem3, mod_rem4)


## DV: arousal rating ####
mod_tst2 <- lm(watch_arou_rating ~ psg_tst, 
               data = df.er_psg %>% filter(intensity == "high" , psg_rem > 0))
summary(mod_tst2)

mod_rem5 <- lm(watch_arou_rating ~ psg_rem, 
               data = df.er_psg %>% filter(intensity == "high" , psg_rem > 0))
summary(mod_rem5)

mod_rem6 <- lm(watch_arou_rating ~ psg_tst + psg_rem, 
               data = df.er_psg %>% filter(intensity == "high"))
summary(mod_rem6)

## interaction effect
mod_rem7 <- lmer(watch_arou_rating ~ psg_rem+intensity_num+ (1|participant), 
               data = df.er_psg)
mod_rem8 <- lmer(watch_arou_rating ~ psg_rem*intensity_num+ (1|participant), 
               data = df.er_psg)
anova(mod_rem7, mod_rem8)
```

```{r analysis: SWS and ER-task linear models}
## DV: valence ratings ####
mod_sws1 <- lm(watch_val_rating ~ psg_sws,
               data = df.er_psg %>% filter(intensity == "high" , psg_rem > 0))
summary(mod_sws1)

mod_sws7 <- lm(watch_val_rating ~ psg_sws + psg_tst,
               data = df.er_psg %>% filter(intensity == "mid" , psg_rem > 0))
summary(mod_sws7)

# interaction
mod_sws2 <- lmer(watch_val_rating ~ psg_sws+intensity_num + (1|participant), 
               data = df.er_psg)
mod_sws3 <- lmer(watch_val_rating ~ psg_sws*intensity_num + (1|participant), 
               data = df.er_psg)
anova(mod_sws2, mod_sws3)

## DV: arousal ratings ####
mod_sws4 <- lm(watch_arou_rating ~ psg_sws,
               data = df.er_psg %>% filter(intensity == "high" , psg_rem > 0))
summary(mod_sws4)

mod_sws8 <- lm(watch_val_rating ~ psg_sws + psg_tst,
               data = df.er_psg %>% filter(intensity == "mid" , psg_rem > 0))
summary(mod_sws8)

# interaction
mod_sws5 <- lmer(watch_arou_rating ~ psg_sws+intensity_num + (1|participant), 
               data = df.er_psg)
mod_sws6 <- lmer(watch_arou_rating ~ psg_sws*intensity_num + (1|participant), 
               data = df.er_psg)
anova(mod_sws5, mod_sws6)
```
```{r analyses: TST and ER-task linear models}
## interaction
mod_tst4 <- lmer(watch_val_rating ~ psg_tst*intensity_num +
                   (1+intensity_num|participant), 
               data = df.er_psg)
# lmerTest::lmer(scale(watch_val_rating) ~ scale(psg_tst)*scale(intensity_num) + (1|participant),
#                data = df.er_psg) %>% summary()

mod_tst4c <- lmer(watch_val_rating ~ psg_tst + intensity_num +
                    (1+intensity_num|participant), 
               data = df.er_psg)
anova(mod_tst4c, mod_tst4)



mod_tst3 <- lmer(watch_arou_rating ~ psg_tst*intensity_num +
                   (1+intensity_num|participant), 
               data = df.er_psg) 
# lmerTest::lmer(scale(watch_arou_rating) ~ scale(psg_tst)*scale(intensity_num) + (1|participant),
#                data = df.er_psg) %>% summary()

mod_tst3c <- lmer(watch_arou_rating ~ psg_tst + intensity_num +
                    (1+intensity_num|participant), 
               data = df.er_psg)
anova(mod_tst3c, mod_tst3)

# plot the model prediction
p_tst_mod1 <- mod_tst4 %>% 
  augment()  %>% 
  clean_names() %>% 
  mutate(intensity = ifelse(intensity_num == 0, "ntr",
                            ifelse(intensity_num == 1, "low",
                                   ifelse(intensity_num == 2, "mid",
                                          "high"))),
         intensity = factor(intensity, levels = c("ntr", "low", "mid", "high"))) %>% 
  ggplot(aes(psg_tst, watch_val_rating, color = as.factor(intensity_num))) +
  geom_point(alpha = .3, color = "grey")+
  geom_point(aes(y = fitted), alpha = .5)+
  geom_smooth(aes(y = fitted), method = "lm", se = F)+
  facet_wrap(.~intensity, nrow = 1)+
  labs(title = "model prediction: TST and valence-reactivity",
       x = "total sleep time")+
  theme(legend.position = "none")

p_tst_mod2 <- mod_tst3 %>% 
  augment()  %>% 
  clean_names() %>% 
  mutate(intensity = ifelse(intensity_num == 0, "ntr",
                            ifelse(intensity_num == 1, "low",
                                   ifelse(intensity_num == 2, "mid",
                                          "high"))),
         intensity = factor(intensity, levels = c("ntr", "low", "mid", "high"))) %>% 
  ggplot(aes(psg_tst, watch_arou_rating, color = as.factor(intensity_num))) +
  geom_point(alpha = .3, color = "grey")+
  geom_point(aes(y = fitted), alpha = .5)+
  geom_smooth(aes(y = fitted), method = "lm", se = F)+
  facet_wrap(.~intensity, nrow = 1)+
  labs(title = "model prediction: TST and arousal-reactivity",
       x = "total sleep time")+
  theme(legend.position = "none")

plot_grid(p_tst_mod1, p_tst_mod2, ncol = 1)
ggsave("plots/tst-reactivity models_random_slope.png", width = 9, height = 6)
```


```{r analysis: REM vs. SWS}
## compare SWS and REM
lm(watch_val_rating ~ psg_rem + psg_sws, 
   data = df.er_psg %>% filter(intensity == "high", psg_rem > 0)) %>% 
  summary()

lm(watch_arou_rating ~ psg_rem + psg_sws, 
   data = df.er_psg %>% filter(intensity == "high", psg_rem > 0)) %>% 
  summary()
```

## Trial-level models
```{r analysis: trial-level model, warning=F}
# construct the trial-level data frame
df.er_trial <- agg_df %>% 
  select(participant, block, imageFile, task, intensity, 
         NegRating.response, NegRating.responseTime,
         ArousalRating.response, ArousalRating.responseTime) %>% 
  mutate_at(vars(NegRating.response, NegRating.responseTime,
                 ArousalRating.response, ArousalRating.responseTime), 
            funs(as.numeric)) %>% 
  rename(val_rating = NegRating.response,
         val_RT = NegRating.responseTime,
         arou_rating = ArousalRating.response,
         arou_RT = ArousalRating.responseTime) %>% 
  filter(!is.na(val_rating), !is.na(arou_rating)) # remove NA data

# merge with the psg data
df.er_trial_psg <- left_join(df.er_trial, df.psg_summary, 
                             by = c("participant" = "pid")) %>%
  left_join(., 
            df.ind_diff %>% filter(Progress == 100)%>% select(partID, age, gender),
            by = c("participant" = "partID")) %>% 
  mutate(age = as.numeric(age),
         stimulus = as.numeric(factor(imageFile)),
         intensity_num = ifelse(intensity == "ntr", 0,
                                ifelse(intensity == "low", 1,
                                       ifelse(intensity == "mid", 2,
                                              ifelse(intensity == "high", 3,
                                                     NA)))))
```

Model comparison results showed that there was a significant TST x intensity interaction for valence ratings in the trial-level model, $\chi^2(1) = 5.59, p = .018$. There was also a significant TST x intensity interaction for arousal ratings in the trial-level model, $\chi^2(1) = 19.82, p < .001$. 
```{r trial-level model: Gaussian family}
# build the lmer model for "watch" trials (reactivity)
# valence rating
mod_tst_lmer1 <- lmer(val_rating ~ intensity_num*psg_tst + 
                        (1|participant)+
                        (1|stimulus), 
                      data = df.er_trial_psg %>% filter(task == "watch"))
# model results in table
# mod_tst_lmer1 %>% tidy() %>% kable(digits = 3)

mod_tst_lmer1c <- lmer(val_rating ~ intensity_num+psg_tst + 
                        (1|participant)+
                        (1|stimulus), 
                      data = df.er_trial_psg %>% filter(task == "watch"))

# trial-level model for valence
anova(mod_tst_lmer1c, mod_tst_lmer1)


# arousal rating
mod_tst_lmer2 <- lmer(arou_rating ~ intensity_num*psg_tst + 
                        (1|participant)+
                        (1|stimulus), 
                      data = df.er_trial_psg %>% filter(task == "watch"))
# model results in table
# mod_tst_lmer2 %>% tidy() %>% kable(digits = 3)

mod_tst_lmer2c <- lmer(arou_rating ~ intensity_num+psg_tst + 
                        (1|participant)+
                        (1|stimulus), 
                      data = df.er_trial_psg %>% filter(task == "watch"))

# trial-level model for arousal
anova(mod_tst_lmer2c, mod_tst_lmer2)
```

The trial-level models predictions are plotted below in comparison to the raw data points.
```{r visualize the trial-level models}
p_tst_lmer1 <- mod_tst_lmer1 %>% 
  augment() %>%
  clean_names() %>% 
  mutate(intensity_num = factor(intensity_num, 
                                labels = c("ntr", "low", "mid", "high"))) %>% 
  ggplot(aes(psg_tst, fitted, color = intensity_num))+
  geom_point(aes(y = val_rating), alpha = .1, color = "grey", 
             position = position_jitter(width = 0, height = .1))+
  geom_point(alpha = .1)+
  geom_smooth(method = "lm")+
  labs(title = "Trial-level model: valence", y = "Valence")+
  facet_wrap(.~intensity_num, nrow = 1)+
  theme(legend.position = "none")

p_tst_lmer2 <- mod_tst_lmer2 %>% 
  augment() %>%
  clean_names() %>% 
  mutate(intensity_num = factor(intensity_num, 
                                labels = c("ntr", "low", "mid", "high"))) %>% 
  ggplot(aes(psg_tst, fitted, color = intensity_num))+
  geom_point(aes(y = arou_rating), alpha = .1, color = "grey", 
             position = position_jitter(width = 0, height = .1))+
  geom_point(alpha = .1)+
  geom_smooth(method = "lm")+
  labs(title = "Trial-level model: arousal", y = "Arousal")+
  facet_wrap(.~intensity_num, nrow = 1)+
  theme(legend.position = "none")

plot_grid(p_tst_lmer1, p_tst_lmer2, ncol = 1)
# ggsave("plots/trial_level_models.png", width = 8, height = 6)
```
```{r trial model: ordered logit}

# factorize valence and arousal rating
df.er_trial_psg <- df.er_trial_psg %>% 
  ungroup() %>% 
  mutate(val_rating_factored = factor(val_rating),
         arou_rating_factored = factor(arou_rating))

# valence
mod_tst1_brm_trial <- brm(val_rating ~ psg_tst*intensity_num +
                      (1|participant)+
                      (1|stimulus), 
                    data = df.er_trial_psg %>% filter(task == "watch"), 
                    family = cumulative,
                    save_all_pars = T,
                    file = "tst_brm_trial1",
                    iter = 2000, warmup = 500, chains = 4, cores = 2)
#mod_tst1_brm_trial %>% plot(N = 4)
mod_tst1_brm_trial %>% 
  augment() %>% 
  view()

# arousal
mod_tst2_brm_trial <- brm(arou_rating ~ psg_tst*intensity_num +
                      (1|participant)+
                      (1|stimulus), 
                    data = df.er_trial_psg %>% filter(task == "watch"), 
                    family = cumulative,
                    save_all_pars = T,
                    file = "tst_brm_trial2",
                    iter = 2000, warmup = 500, chains = 4, cores = 2)
#mod_tst2_brm_trial %>% plot(N = 4)
```

```{r visualize the trial-level clmm models}
ggplot(df.er_trial_psg, aes(psg_tst, val_rating))+
  geom_bar(aes(fill = val_rating_factored),
           stat = "identity")+
  # geom_area(fill = "#eaac9e",
  #           color = "black",
  #           size = 0.2)+
  # geom_ribbon(aes(ymin = val_rating, ymax = 1),
  #             fill = "#92bbd1",
  #             color = "black",
  #             size = 0.2)+
  facet_wrap(.~ intensity)


p_tst1_brm_trial <- mod_tst1_brm_trial %>% 
  broom.mixed::augment() %>%
  clean_names() %>% 
  mutate(intensity_num = factor(intensity_num, 
                                labels = c("ntr", "low", "mid", "high"))) %>% 
  ggplot(aes(psg_tst, fitted, color = intensity_num))+
  geom_point(aes(y = val_rating), alpha = .1, color = "grey", 
             position = position_jitter(width = 0, height = .1))+
  geom_point(alpha = .1)+
  geom_smooth(method = "lm")+
  labs(title = "Trial-level model: valence", y = "Valence")+
  facet_wrap(.~intensity_num, nrow = 1)+
  theme(legend.position = "none")

plot_grid(p_tst_lmer1, p_tst_lmer2, ncol = 1)
# ggsave("plots/trial_level_models.png", width = 8, height = 6)
```



### Bayesian models

We also built the same models using Bayesian package `brms` and compute the Bayes Factor for the effect of REM and SWS. First, we built the bayesian models with TST x intensity for both valence and arousal. The trace plots showed that the Monte Carlo sampling was effective.
```{r TST model using Bayesian brms}
# build mod_tst1 using brms (weakly-informative prior)
mod_tst1_brm <- brm(valence ~ psg_tst*intensity_num + (1|participant), 
               data = df.er_psg,
               family = gaussian,
               save_all_pars = T,
               file = "tst1_brm",
               iter = 1000, warmup = 500, chains = 4, cores = 2)
mod_tst1_brm %>% plot(N = 6)

# build mod_tst2 using brms (weakly-informative prior)
mod_tst2_brm <- brm(arousal ~ psg_tst*intensity_num + (1|participant), 
               data = df.er_psg,
               family = gaussian,
               save_all_pars = T,
               file = "tst2_brm",
               iter = 1000, warmup = 500, chains = 4, cores = 2)
mod_tst2_brm %>% plot(N = 6)
```

Second, we built the bayesian models with TST x intensity plus REM duration for both valence and arousal. The trace plots showed that the Monte Carlo sampling was effective. Following that, we computed the Bayes Factor of the REM effect.
```{r REM model using Bayesian brms}
# build mod_rem1 using brms (weakly-informative prior)
mod_rem1_brm <- brm(valence ~ psg_tst*intensity_num + psg_rem + (1|participant), 
               data = df.er_psg,
               family = gaussian,
               save_all_pars = T,
               file = "rem1_brm",
               iter = 1000, warmup = 500, chains = 4, cores = 2)
# show the posterior
# print(mod_rem1_brm)
# the trace plots
plot(mod_rem1_brm, N = 7)

# mod_rem1_brm %>%
#   posterior_samples() %>%
#   select(starts_with("b_psg_rem")) %>%
#   gather("variable", "value") %>%
#   ggplot(aes(y = variable, x = value))+
#   tidybayes::geom_halfeyeh()+
#   geom_vline(xintercept = 0, linetype = 2)+
#   labs(title = "Model on valence")
# ggsave("../../figures/rem_valence_posterior.png", width = 4, height = 3)


# build mod_rem2 using brms (weakly-informative prior)
mod_rem2_brm <- brm(arousal ~ psg_tst*intensity_num + psg_rem + (1|participant), 
               data = df.er_psg,
               family = gaussian,
               save_all_pars = T,
               file = "rem2_brm",
               iter = 1000, warmup = 500, chains = 4, cores = 2)
# print the model results
# print(mod_rem2_brm)
# the trace plot
plot(mod_rem2_brm, N = 7)
```

```{r Bayes factor: REM, message=F, results='hide'}
# compare the rem_model with the tst_model
# for valence
BF_rem1 = bayes_factor(mod_rem1_brm, mod_tst1_brm)
# for arousal
BF_rem2 = bayes_factor(mod_rem2_brm, mod_tst2_brm)

```
```{r Bayes factor: REM2}
# pring the BFs
BF_rem1
BF_rem2
```

For valence, the Bayes Factor of REM is `r BF_rem1[1] %>% as.numeric() %>% round(4)`. For arousal, the Bayes Factor of REM is `r BF_rem2[1] %>% as.numeric() %>% round(4)`. Both results provided evidence for a null effect of REM on emotional reactivity.

Third, we built the bayesian models with TST x intensity plus SWS duration for both valence and arousal. The trace plots showed that the Monte Carlo sampling was effective. Following that, we computed the Bayes Factor of the SWS effect.
```{r SWS model using Bayesian brms}
# build mod_sws1 using brms (weakly-informative prior)
mod_sws1_brm <- brm(valence ~ psg_tst*intensity_num + psg_sws + (1|participant), 
               data = df.er_psg,
               family = gaussian,
               save_all_pars = T,
               file = "sws1_brm",
               iter = 1000, warmup = 500, chains = 4, cores = 2)
# # show the posterior
# print(mod_sws1_brm)
# # the trace plots
plot(mod_sws1_brm, N = 7)

# build mod_sws2 using brms (weakly-informative prior)
mod_sws2_brm <- brm(arousal ~ psg_tst*intensity_num + psg_sws + (1|participant), 
               data = df.er_psg,
               family = gaussian,
               save_all_pars = T,
               file = "sws2_brm",
               iter = 1000, warmup = 500, chains = 4, cores = 2)
# # print the model results
# print(mod_sws2_brm)
# # the trace plot
plot(mod_sws2_brm, N = 7)
```

```{r Bayes factor: SWS, message=F, results='hide'}
# compare the rem_model with the tst_model
# for valence
BF_sws1 = bayes_factor(mod_sws1_brm, mod_tst1_brm)
BF_sws1
# for arousal
BF_sws2 = bayes_factor(mod_sws2_brm, mod_tst2_brm)
BF_sws2
```
```{r Bayes factor: SWS2}
# pring the BFs
BF_sws1
BF_sws2
```
For arousal, the Bayes Factor of SWS is `r BF_sws1[1] %>% as.numeric() %>% round(4)`. For arousal, the Bayes Factor of SWS is `r BF_sws2[1] %>% as.numeric() %>% round(4)`. Both results provided evidence for a null effect of SWS on emotional reactivity.

Finally, we built the bayesian models for understanding the effects of all sleep stages at the same time. All variables are standardized before fitting the model so that the parameter posteriors are in the same scale. The trace plots showed that the Monte Carlo sampling was effective. The model results are presented in the following tables.
```{r sleep stages model using Bayesian brms, warning=F}
# build mod_sleep1 using brms
mod_sleep1_brm = brm(scale(valence)~scale(psg_n1)+scale(psg_n2)+
                       scale(psg_sws)+scale(psg_rem)+
                       scale(intensity_num)+(1|participant),
                     data = df.er_psg,
                     family = gaussian,
                     file = "sleep1_brm",
                     iter = 1000, warmup = 500, chains = 4, cores = 2)

plot(mod_sleep1_brm, N = 8)
# print the table
tidy(mod_sleep1_brm) %>% 
  filter(grepl("b_", term)) %>% 
  kable(digits = 3)

# build mod_sleep2 using brms
mod_sleep2_brm = brm(scale(arousal) ~ ~scale(psg_n1)+scale(psg_n2)+
                       scale(psg_sws)+scale(psg_rem)+
                       scale(intensity_num)+(1|participant),
                     data = df.er_psg,
                     family = gaussian,
                     file = "sleep2_brm",
                     iter = 1000, warmup = 500, chains = 4, cores = 2)

plot(mod_sleep2_brm, N = 8)
# print the table
tidy(mod_sleep2_brm) %>% 
  filter(grepl("b_", term)) %>% 
  kable(digits = 3)

```

The posterior distributions of the effects of each sleep stage on valence and arousal are presented in the figure below. Consistent with the results of the non-Bayesian models, Stage 2 sleep seemed to have the largest effect on emotional ractivity among all sleep stages.
```{r visualize the posterior of sleep stages, warning=F}
# plot the standardized b for each predictor
p_sleep1_brm <- mod_sleep1_brm %>%
  posterior_samples() %>%
  select(starts_with("b_scale")) %>%
  gather("variable", "value") %>%
  ggplot(aes(y = variable, x = value))+
  tidybayes::geom_halfeyeh()+
  geom_vline(xintercept = 0, linetype = 2)+
  labs(title = "Model on valence")


# plot the standardized b for each predictor
p_sleep2_brm <- mod_sleep2_brm %>%
  posterior_samples() %>%
  select(starts_with("b_scale")) %>%
  gather("variable", "value") %>%
  ggplot(aes(y = variable, x = value))+
  tidybayes::geom_halfeyeh()+
  geom_vline(xintercept = 0, linetype = 2)+
  labs(title = "Model on arousal")

plot_grid(p_sleep1_brm, p_sleep2_brm, ncol = 1)
# ggsave("../../figures/sleep_stages_posterior.png", width = 8, height = 10)
```


```{r simulation}
# simulation dataframe
df.simul <- df.er_psg %>% 
  filter(!is.na(psg_tst), intensity == "high") 


fun.simulate <- function(){
  # reactivity 1: only related to TST
  df.simul$reactivity1 <- df.simul$psg_tst*0.01 + rnorm(41, 0, 1)
  # reactivity 2: related to TST and REM
  df.simul$reactivity2 <- df.simul$psg_tst*0.008 + df.simul$psg_rem*0.01 + rnorm(41, 0, 1)
  
  # ggpairs(df.simul %>% select(psg_tst, psg_rem, psg_rem_perc, reactivity1, reactivity2))
  # ggsave("plots/simulation_correlation.png", width = 8, height = 7)

  m1 <- lm(reactivity1 ~ psg_tst + psg_rem, data = df.simul) %>% tidy()
  m2 <- lm(reactivity1 ~ psg_rem_perc, data = df.simul) %>% tidy()
  m3 <- lm(reactivity2 ~ psg_tst + psg_rem, data = df.simul) %>% tidy()
  m4 <- lm(reactivity2 ~ psg_rem_perc, data = df.simul) %>% tidy()
  
  return(c(m1$p.value[3], m2$p.value[2], m3$p.value[3], m4$p.value[2]))
}

# simulate 1k times
sim_result <- replicate(10000, fun.simulate())


# construc the data fame from the simulated results
sim_result <- t(sim_result) %>% as.tibble()
names(sim_result) <- c("p1_reactivity1", "p2_reactivity1", 
                         "p1_reactivity2", "p2_reactivity2")
sim_result_t <- sim_result %>% 
  gather(p_type, p_value, p1_reactivity1:p2_reactivity2) %>% 
  separate(col = p_type, into = c("p_type", "DV"), sep = "_")

sim_result_t$p_type <- factor(sim_result_t$p_type, levels = c("p2", "p1"))

# summarize p-values
sim_result_t %>% 
  group_by(p_type, DV) %>% 
  summarise(number = sum(p_value < .05),
            rej_rate = number/n()) %>% 
  arrange(DV)



ggplot(sim_result_t, 
       aes(x = p_value))+
  geom_density(aes(fill = p_type), alpha = .3)+
  geom_text(data = data.frame(DV = c("reactivity1", "reactivity2"),
                       text_x = c(.5, .5),
                       text_y = c(5 ,5),
                       text = c("null effect data:\np1 (using tst+rem) rejection rate = .0478\np2 (using rem%) rejection rate = .0343",
                                "REM effect data:\np1 (using tst+rem) rejection rate = .4251\n p2 (using rem%) rejection rate = .4695"),
                       p_type = c("1","2")
                       ),
            aes(text_x, text_y,
                label = text))+
  facet_wrap(DV ~ ., ncol = 1)+
  geom_vline(xintercept = .05, 
             linetype = 2)

ggsave("plots/simulation_distribution.png", width = 8, height = 6)
```

```{r check ERP data}
library(readxl)
df.LPP <- read_excel('C:/Users/tepzh/Desktop/R56 EEG/28-Feb-2019-trial-LPP-export.xlsx')
df.LPP$task <- factor(df.LPP$task, levels = c("Watch", "Rethink", "Distract"))
df.LPP$intensity <- factor(df.LPP$assign, levels = c("ntr", "low", "mid", "high"))
df.LPP$Subject <- factor(df.LPP$Subject)


df.LPP_summary_by_task <- df.LPP %>% 
  group_by(Subject, intensity) %>% 
                summarise(LPP_early_P3PzP4 = mean(LPP_early_P3PzP4),
                          LPP_mid_P3PzP4 = mean(LPP_mid_P3PzP4),
                          LPP_late_P3PzP4 = mean(LPP_late_P3PzP4))

# can gather the data and plot once here!
# early-LPP
p_LPP_early <- ggplot(df.LPP, # %>% filter(intensity %in% c("ntr", "high")),
       aes(intensity, 
           LPP_early_P3PzP4, 
           group = Subject, 
           color = Subject))+
  geom_violin(data = df.LPP_summary_by_task,
              aes(group = NULL),
              color = "black",
              fill = "white")+
  stat_summary(fun.y = "mean",
               geom = "line",
               alpha = .3)+
  stat_summary(fun.y = "mean",
               geom = "point",
               alpha = .3)+
  stat_summary(aes(group = NULL),
               size = 4,
               alpha = .5,
               color = "black",
               fun.y = "mean",
               geom = "point")+
  stat_summary(aes(group = NULL),
               size = 1,
               alpha = .5,
               color = "black",
               fun.data = "mean_cl_boot",
               geom = "linerange")+
  labs(title = "early LPP (300 - 500 ms)")+
  #ylim(-0.010, 0.005)+
  theme(legend.position = "none")


# mid-LPP
p_LPP_mid <-ggplot(df.LPP, # %>% filter(intensity %in% c("ntr", "high")),
       aes(intensity, 
           LPP_mid_P3PzP4, 
           group = Subject, 
           color = Subject))+
  geom_violin(data = df.LPP_summary_by_task,
              aes(group = NULL),
              color = "black",
              fill = "white")+
  stat_summary(fun.y = "mean",
               geom = "line",
               alpha = .3)+
  stat_summary(fun.y = "mean",
               geom = "point",
               alpha = .3)+
  stat_summary(aes(group = NULL),
               size = 4,
               alpha = .5,
               color = "black",
               fun.y = "mean",
               geom = "point")+
  stat_summary(aes(group = NULL),
               size = 1,
               alpha = .5,
               color = "black",
               fun.data = "mean_cl_boot",
               geom = "linerange")+
  labs(title = "mid LPP (500 - 1000 ms)")+
  #ylim(-0.010, 0.005)+
  theme(legend.position = "none")

# late-LPP
p_LPP_late <-ggplot(df.LPP, # %>% filter(intensity %in% c("ntr", "high")),
       aes(intensity, 
           LPP_late_P3PzP4, 
           group = Subject, 
           color = Subject))+
  geom_violin(data = df.LPP_summary_by_task,
              aes(group = NULL),
              color = "black",
              fill = "white")+
  stat_summary(fun.y = "mean",
               geom = "line",
               alpha = .3)+
  stat_summary(fun.y = "mean",
               geom = "point",
               alpha = .3)+
  stat_summary(aes(group = NULL),
               size = 4,
               alpha = .5,
               color = "black",
               fun.y = "mean",
               geom = "point")+
  stat_summary(aes(group = NULL),
               size = 1,
               alpha = .5,
               color = "black",
               fun.data = "mean_cl_boot",
               geom = "linerange")+
  labs(title = "late LPP (1000 - 5000 ms)")+
  #ylim(-0.010, 0.005)+
  theme(legend.position = "none")

plot_grid(p_LPP_early, p_LPP_mid, p_LPP_late, ncol = 2)
ggsave("plots/LPP amplitudes.png", width = 8, height = 6)
```

```{r Check EMG data}
emg_path = "Z:/SBER/SBER1_physio_R56/data/Full_Study/session2_data/results/emg_results"
emg_files <- list.files(emg_path, full.names = T)
emg_files <- emg_files[grep("SBER.EMG_Corr.[0-9]{4}.*txt$", emg_files)]

# extract the participant numbers
emg_pids = str_sub(emg_files, str_length(emg_files) - 7, str_length(emg_files) - 4)

# read the datafiles and save to a list
list.emg <- map(emg_files, ~ read.csv(file = .x, header = F, sep = "\t"))

#view(list.emg[[1]])
# create the combind dataframe
df.emg <- bind_rows(list.emg[1:66])
# add variable names
names(df.emg) <- c("participant", "task", "trial", "time_segment", "emg_corr")
df.emg$participant <- factor(df.emg$participant)
df.emg$task <- factor(df.emg$task, levels = c("Watch", "Rethink", "Distract"))


# check the distribution of data
emg.quantile = quantile(df.emg$emg_corr, probs = seq(0, 1, .01))
plot(emg.quantile)
title("EMG quantile")
abline(h = 20, lty = 2)
text(x = 90, y = 25, "EMG = 20")

ggplot(df.emg, aes(participant, emg_corr))+
  geom_point(alpha = .2, 
             shape = 20, 
             # size = 2,
             color = "blue")+
  stat_summary(fun.y = mean,
               geom = "point",
               color = "red",
               alpha = .5)+
  labs("EMG raw data points for each participant")+
  theme(axis.text.x = element_text(hjust = 1, angle = 45, 
                                   size = 5, color = "black"))

ggsave("plots/emg_raw_points.png", width = 8, height = 6)



# visualize the EMG 
df.emg_summary = df.emg %>% 
                group_by(participant, task) %>% 
                summarise(emg_corr = mean(emg_corr))

ggplot(df.emg, aes(task, emg_corr, 
                   group = participant, color = participant))+
  geom_violin(df.emg_summary,
              aes(group = NULL),
              color = "black",
              fill = "white")+
  stat_summary(fun.y = "mean",
               geom = "line",
               alpha = .3)+
  stat_summary(fun.y = "mean",
               geom = "point",
               alpha = .3)+
  stat_summary(aes(group = NULL),
               size = 4,
               alpha = .5,
               color = "black",
               fun.y = "mean",
               geom = "point")+
  # stat_summary(aes(group = NULL),
  #              size = 1,
  #              alpha = .5,
  #              color = "black",
  #              fun.data = "mean_cl_boot",
  #              geom = "linerange")+
  labs(title = "Corrugator EMG (averaged across trials)")+
  theme(legend.position = "none")

ggsave("plots/emg_distribution.png", width = 8, height = 6)


ggplot(df.emg, aes(time_segment, emg_corr,
                   group = participant, color = participant))+
  stat_summary(fun.y = "mean",
               geom = "line",
               alpha = .3)+
  stat_summary(fun.y = "mean",
               geom = "point",
               alpha = .3)+
  stat_summary(aes(group = NULL),
               size = 2,
               alpha = .5,
               color = "black",
               fun.y = "mean",
               geom = "point")+
  stat_summary(aes(group = NULL),
               size = 1,
               alpha = .5,
               color = "black",
               fun.data = "mean_cl_boot",
               geom = "linerange")+
  stat_summary(aes(group = NULL),
               size = 1,
               alpha = .5,
               color = "black",
               fun.y = "mean",
               geom = "line")+
  labs(title = "Time change of EMG in a trial", x = "time (s)")+
  xlim(0, 6)+
  theme(legend.position = "none")

ggsave("plots/emg_time_change.png", width = 8, height = 6)
  


df.emg_summary %>% filter(emg_corr > 10)
df.emg_summary %>% filter(participant == 1581)
```


```{r save workspace}
save.image('C:/Users/tepzh/Dropbox/Stanford/FYP/fyp_study2/data/workspace_fyp_study2.Rdata')

```


